---
title: "distill and timetk"
description: |
  today's directed wanderings
author: Brandon Farr
date: 2021-01-26
categories:
  - timetk
  - distill
output:
  distill::distill_article:
    self_contained: false
draft: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Today, I worked on the following:

1. setting up this [blog](https://orderlyquant.github.io/wilting)
2. running through the weekly Free R Tip from [business-science](https://www.business-science.io/)
3. read through [GIT SQL Together](https://caitlinhudon.com/2018/11/28/git-sql-together/)

## Distill blog

Set up this blog using the `distill` package and [github pages](github.pages.io).
A few takeaways:

- see [here](https://github.blog/2016-12-09-publishing-with-github-pages-now-as-easy-as-1-2-3/) for details on getting your repo right
- see [About](https://orderlyquant.github.io/wilting/about.html) for other details

## Free R Tip

This week's tip was about many ways to create data frames/tibbles. I was familiar
with these, but found the use of `timetk::tk_make_timeseries` informative.

```{r, echo=TRUE}

tibble::tibble(
    date          = timetk::tk_make_timeseries("2010", length_out = 12, by = "quarter"),
    interest_rate = (seq(12, 3, length.out = 12) * (sin(1:12) + 2)) / 12
)

```

Note that the `by` argument is very flexible:

- `by` - a character string, containing one of "sec", "min", "hour", "day", "week",
  "month", "quarter" or "year". You can create regularly spaced sequences using
  phrases like by = "10 min"


## GIT + SQL

1. You will always need that query again, so "GIT" it
2. Queries change through time, you *may* need a previous version when you realize
   the *improved* query "broke" something in a way that wasn't immediately obvious
   - save with a `.sql` extension

Looks like there are other interesting posts from this site. I added one to Notion,
so it may be covered here someday.
