---
title: "Ha`rvest`ing S&P 500 Sectors"
description: |
  Using open source data to look at within industry correlations.
author:
  - name: Brandon Farr
    url: {}
date: 01-28-2021
categories:
  - tidyquant
  - rvest
  - janitor
  - finance
output:
  distill::distill_article:
    self_contained: false
draft: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(tidyverse)  # for tidy/dplyr work
library(rvest)      # for web-scraping

```

A little work with the `rvest` package. Try to get the current S&P 500
constituents from Wikipedia.

```{r}

wiki_sp500 <- read_html(
  "https://en.wikipedia.org/wiki/List_of_S%26P_500_companies"
)

```

According to my [SelectorGadget](https://rvest.tidyverse.org/articles/articles/selectorgadget.html)
work, the element I am looking for is "#constituents"

```{r}

wiki_sp500 %>%
  html_element("#constituents")

```

Since this is a table, it needs to be passed through the `html_table`
function.

```{r}

sp500_tbl <- wiki_sp500 %>%
  html_element("#constituents") %>%
  html_table() %>%
  janitor::clean_names()

sp500_tbl

```

Wow! Sure enough, we get a tibble with what we want. Notice the use of
`janitor` to clean up the column names for convenience. *I think I should
always do this.*

Let's look at all of the stocks in the same `gics_sub_industry` as Apple.

```{r}

aapl_sub_ind <- sp500_tbl %>%
  filter(symbol == "AAPL") %>%
  pull(gics_sub_industry)

aapl_peers_tbl <- sp500_tbl %>%
  filter(gics_sub_industry == aapl_sub_ind) %>%
  select(symbol, security, matches("^gics"))

aapl_peers_tbl

```

This is a great list to pass to `tidyquant` and try pulling some pricing data.

```{r}

library(tidyquant)

aapl_peers_pricing <- aapl_peers_tbl %>%
  pull(symbol) %>%
  tq_get(get = "tiingo")

```

By default, we get about a years worth of prices. Let's calculate returns:

```{r}

aapl_peers_returns <- aapl_peers_pricing %>%
  group_by(symbol) %>%
  tq_transmute(adjusted, periodReturn, period = "daily", col_rename = "return")

```

Now, calculate the correlation matrix:

```{r}

aapl_peers_returns %>%
  
  # put returns into columns
  pivot_wider(names_from = symbol, values_from = return) %>%
  
  # first day return is 0, so remove row
  slice(-1) %>%
  
  # remove date column, so you can call `cor` on entire tibble
  select(-date) %>%
  
  # calculate correlations
  cor() %>%
  
  # `cor` returns a matrix, convert back into tibble
  as_tibble(rownames = "symbol_1") %>%
  
  # transform into tidy format for easier processing
  pivot_longer(-1, names_to = "symbol_2", values_to = "cor") %>%
  
  # remove the correlations of a stock with itself
  filter(!(symbol_1 == symbol_2)) %>%
  
  # group by symbol_1 and arrange descending by correlation
  group_by(symbol_1) %>%
  arrange(desc(cor)) %>%
  
  # look at first row in each group == highest correlated stock
  slice(1)

```


